Arrows for Parallel Computations

### 2017-05-19 paper/content/main.tex
formatting of topo skels

\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{ring}\mathbin{::}(\Conid{ArrowLoop}\;\Varid{arr},\Conid{Future}\;\Varid{fut}\;\Varid{r},\Conid{ArrowParallel}\;\Varid{arr}\;(\Varid{i},\Varid{fut}\;\Varid{r})\;(\Varid{o},\Varid{fut}\;\Varid{r})\;\Varid{conf})\Rightarrow {}\<[E]%
\\
\>[B]{}\hsindent{5}{}\<[5]%
\>[5]{}\Varid{conf}\to \Varid{arr}\;(\Varid{i},\Varid{r})\;(\Varid{o},\Varid{r})\to \Varid{arr}\;[\mskip1.5mu \Varid{i}\mskip1.5mu]\;[\mskip1.5mu \Varid{o}\mskip1.5mu]{}\<[E]%
\\
\>[B]{}\Varid{ring}\;\Varid{conf}\;\Varid{f}\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{loop}\;(\Varid{second}\;(\Varid{rightRotate}\mathbin{>\!\!>\!\!>}\Varid{lazy})\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{arr}\;(\Varid{uncurry}\;\Varid{zip})\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{parMap}\;\Varid{conf}\;(\Varid{second}\;\Varid{get}\mathbin{>\!\!>\!\!>}\Varid{f}\mathbin{>\!\!>\!\!>}\Varid{second}\;\Varid{put})\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{arr}\;\Varid{unzip}){}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{rightRotate}\mathbin{::}(\Conid{Arrow}\;\Varid{arr})\Rightarrow \Varid{arr}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
\\
\>[B]{}\Varid{rightRotate}\mathrel{=}\Varid{arr}\mathbin{\$}\lambda \Varid{list}\to \mathbf{case}\;\Varid{list}\;\mathbf{of}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}[\mskip1.5mu \mskip1.5mu]\to [\mskip1.5mu \mskip1.5mu]{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{xs}\to \Varid{last}\;\Varid{xs}\mathbin{:}\Varid{init}\;\Varid{xs}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{lazy}\mathbin{::}(\Conid{Arrow}\;\Varid{arr})\Rightarrow \Varid{arr}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
\\
\>[B]{}\Varid{lazy}\mathrel{=}\Varid{arr}\;(\lambda \mathord{\sim}(\Varid{x}\mathbin{:}\Varid{xs})\to \Varid{x}\mathbin{:}\Varid{lazy}\;\Varid{xs}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks
\caption{Final definition of the \ensuremath{\Varid{ring}} skeleton.}
\label{fig:ringFinal}
\end{figure}
 % $ %% formatting
% and lazy:
% \begin{code}
% lazy :: (Arrow arr) => arr [a] [a]
% lazy = arr (\ ~(x:xs) -> x : lazy xs
% \end{code}
%% we have shown this already.

\subsection{Torus skeleton}
\begin{figure}
	\includegraphics[scale=0.75]{images/torus}
	\caption{Schematic depiction of the torus skeleton}
	\label{fig:ringTorusImg}
\end{figure}
If we take the concept of a ring from \ref{sec:ring} one dimension further, we get a torus (Fig.~\ref{fig:ringTorusImg},~\ref{fig:torus}). Every node sends ands receives data from horizontal and vertical neighbours in each communication round.
With our Parallel Arrows we re-implement the torus combinator\footnote{Available on Hackage: \url{https://hackage.haskell.org/package/edenskel-2.0.0.2/docs/Control-Parallel-Eden-Topology.html}.} from Eden---yet again with the help of the \ensuremath{\Conid{ArrowLoop}} typeclass.

@@

@@

@@

\>[B]{}\Varid{torus}\mathbin{::}(\Conid{ArrowLoop}\;\Varid{arr},\Conid{ArrowChoice}\;\Varid{arr},\Conid{ArrowApply}\;\Varid{arr},{}\<[E]%

@@

\>[9]{}\Conid{ArrowParallel}\;\Varid{arr}\;(\Varid{c},\Varid{fut}\;\Varid{a},\Varid{fut}\;\Varid{b})\;(\Varid{d},\Varid{fut}\;\Varid{a},\Varid{fut}\;\Varid{b})\;\Varid{conf},{}\<[E]%
\>[9]{}\Conid{Future}\;\Varid{fut}\;\Varid{a},\Conid{Future}\;\Varid{fut}\;\Varid{b})\Rightarrow {}\<[E]%

@@

\>[9]{}\Varid{loop}\;(\Varid{second}\;((\Varid{mapArr}\;\Varid{rightRotate}\mathbin{>\!\!>\!\!>}\Varid{lazy})\mathbin{*\!*\!*}{}\<[E]%
\>[17]{}(\Varid{arr}\;\Varid{rightRotate}\mathbin{>\!\!>\!\!>}\Varid{lazy}))\mathbin{>\!\!>\!\!>}{}\<[E]%

@@

\>[9]{}(\Varid{arr}\;\Varid{length}\mathbin{>\!\!>\!\!>}\Varid{arr}\;\Varid{unshuffle})\mathbin{\&\!\&\!\&}{}\<[E]%
\>[17]{}(\Varid{shuffle}\mathbin{>\!\!>\!\!>}\Varid{parMap}\;\Varid{conf}\;(\Varid{ptorus}\;\Varid{f}))\mathbin{>\!\!>\!\!>}{}\<[E]%
\>[9]{}\Varid{app}\mathbin{>\!\!>\!\!>}{}\<[E]%

@@

\>[9]{}\Varid{arr}\;(\Varid{c},\Varid{a},\Varid{b})\;(\Varid{d},\Varid{a},\Varid{b})\to {}\<[E]%

@@

\>[9]{}\Varid{arr}\;(\Varid{c},\Varid{fut}\;\Varid{a},\Varid{fut}\;\Varid{b})\;(\Varid{d},\Varid{fut}\;\Varid{a},\Varid{fut}\;\Varid{b}){}\<[E]%
\>[B]{}\Varid{ptorus}\;\Varid{f}\mathrel{=}{}\<[E]%
\>[9]{}\Varid{arr}\;(\lambda \mathord{\sim}(\Varid{c},\Varid{a},\Varid{b})\to (\Varid{c},\Varid{get}\;\Varid{a},\Varid{get}\;\Varid{b}))\mathbin{>\!\!>\!\!>}\Varid{f}\mathbin{>\!\!>\!\!>}{}\<[E]%
\>[9]{}\Varid{arr}\;(\lambda \mathord{\sim}(\Varid{d},\Varid{a},\Varid{b})\to (\Varid{d},\Varid{put}\;\Varid{a},\Varid{put}\;\Varid{b})){}\<[E]%

@@

\column{25}{@{}>{\hspre}l<{\hspost}@{}}%

@@

\>[B]{}\Varid{nodefunction}\mathbin{::}\Conid{Int}\to {}\<[E]%

@@

\>[9]{}((\Conid{Matrix},\Conid{Matrix}),[\mskip1.5mu \Conid{Matrix}\mskip1.5mu],[\mskip1.5mu \Conid{Matrix}\mskip1.5mu])\to {}\<[E]%

@@

\>[9]{}([\mskip1.5mu \Conid{Matrix}\mskip1.5mu],[\mskip1.5mu \Conid{Matrix}\mskip1.5mu],[\mskip1.5mu \Conid{Matrix}\mskip1.5mu]){}\<[E]%

@@

\>[B]{}\Varid{nodefunction}\;\Varid{n}\;((\Varid{bA},\Varid{bB}),\Varid{rows},\Varid{cols})\mathrel{=}{}\<[E]%

@@

\>[9]{}([\mskip1.5mu \Varid{bSum}\mskip1.5mu],\Varid{bA}\mathbin{:}\Varid{nextAs},\Varid{bB}\mathbin{:}\Varid{nextBs}){}\<[E]%
\>[17]{}\mathbf{where}\;\Varid{bSum}\mathrel{=}{}\<[E]%
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{foldl'}\;\Varid{matAdd}\;(\Varid{matMult}\;\Varid{bA}\;\Varid{bB})\;(\Varid{zipWith}\;\Varid{matMult}\;\Varid{nextAs}\;\Varid{nextBs}){}\<[E]%
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{nextAs}\mathrel{=}\Varid{take}\;(\Varid{n}\mathbin{-}\mathrm{1})\;\Varid{rows}{}\<[E]%
\>[17]{}\hsindent{8}{}\<[25]%
\>[25]{}\Varid{nextBs}\mathrel{=}\Varid{take}\;(\Varid{n}\mathbin{-}\mathrm{1})\;\Varid{cols}{}\<[E]%

>>>>

\begin{figure}[h]
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{ring}\mathbin{::}(\Conid{ArrowLoop}\;\Varid{arr},\Conid{Future}\;\Varid{fut}\;\Varid{r},\Conid{ArrowParallel}\;\Varid{arr}\;(\Varid{i},\Varid{fut}\;\Varid{r})\;(\Varid{o},\Varid{fut}\;\Varid{r})\;\Varid{conf})\Rightarrow {}\<[E]%
\\
\>[B]{}\hsindent{5}{}\<[5]%
\>[5]{}\Varid{conf}\to \Varid{arr}\;(\Varid{i},\Varid{r})\;(\Varid{o},\Varid{r})\to \Varid{arr}\;[\mskip1.5mu \Varid{i}\mskip1.5mu]\;[\mskip1.5mu \Varid{o}\mskip1.5mu]{}\<[E]%
\\
\>[B]{}\Varid{ring}\;\Varid{conf}\;\Varid{f}\mathrel{=}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{loop}\;(\Varid{second}\;(\Varid{rightRotate}\mathbin{>\!\!>\!\!>}\Varid{lazy})\mathbin{>\!\!>\!\!>}\Varid{arr}\;(\Varid{uncurry}\;\Varid{zip})\mathbin{>\!\!>\!\!>}{}\<[E]%
\\
\>[B]{}\hsindent{9}{}\<[9]%
\>[9]{}\Varid{parMap}\;\Varid{conf}\;(\Varid{second}\;\Varid{get}\mathbin{>\!\!>\!\!>}\Varid{f}\mathbin{>\!\!>\!\!>}\Varid{second}\;\Varid{put})\mathbin{>\!\!>\!\!>}\Varid{arr}\;\Varid{unzip}){}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{rightRotate}\mathbin{::}(\Conid{Arrow}\;\Varid{arr})\Rightarrow \Varid{arr}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
\\
\>[B]{}\Varid{rightRotate}\mathrel{=}\Varid{arr}\mathbin{\$}\lambda \Varid{list}\to \mathbf{case}\;\Varid{list}\;\mathbf{of}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}[\mskip1.5mu \mskip1.5mu]\to [\mskip1.5mu \mskip1.5mu]{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{xs}\to \Varid{last}\;\Varid{xs}\mathbin{:}\Varid{init}\;\Varid{xs}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{lazy}\mathbin{::}(\Conid{Arrow}\;\Varid{arr})\Rightarrow \Varid{arr}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
\\
\>[B]{}\Varid{lazy}\mathrel{=}\Varid{arr}\;(\lambda \mathord{\sim}(\Varid{x}\mathbin{:}\Varid{xs})\to \Varid{x}\mathbin{:}\Varid{lazy}\;\Varid{xs}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks
\caption{Final definition of the \ensuremath{\Varid{ring}} skeleton.}
\label{fig:ringFinal}
\end{figure}
 % $ %% formatting
% and lazy:
% \begin{code}
% lazy :: (Arrow arr) => arr [a] [a]
% lazy = arr (\ ~(x:xs) -> x : lazy xs
% \end{code}
%% we have shown this already.

\subsection{Torus skeleton}
\begin{figure}
	\includegraphics[scale=0.75]{images/torus}
	\caption{Schematic depiction of the \ensuremath{\Varid{torus}} skeleton.}
	\label{fig:ringTorusImg}
\end{figure}
If we take the concept of a ring from \ref{sec:ring} one dimension further, we get a torus (Fig.~\ref{fig:ringTorusImg},~\ref{fig:torus}). Every node sends ands receives data from horizontal and vertical neighbours in each communication round.
With our Parallel Arrows we re-implement the \ensuremath{\Varid{torus}} combinator\footnote{Available on Hackage: \url{https://hackage.haskell.org/package/edenskel-2.0.0.2/docs/Control-Parallel-Eden-Topology.html}.} from Eden---yet again with the help of the \ensuremath{\Conid{ArrowLoop}} typeclass.

@@

\olcomment{swapped type classes}

@@

\column{11}{@{}>{\hspre}l<{\hspost}@{}}%

@@

\>[B]{}\Varid{torus}\mathbin{::}(\Conid{ArrowLoop}\;\Varid{arr},\Conid{ArrowChoice}\;\Varid{arr},\Conid{ArrowApply}\;\Varid{arr},\Conid{Future}\;\Varid{fut}\;\Varid{a},\Conid{Future}\;\Varid{fut}\;\Varid{b},{}\<[E]%

@@

\>[B]{}\hsindent{11}{}\<[11]%
\>[11]{}\Conid{ArrowParallel}\;\Varid{arr}\;(\Varid{c},\Varid{fut}\;\Varid{a},\Varid{fut}\;\Varid{b})\;(\Varid{d},\Varid{fut}\;\Varid{a},\Varid{fut}\;\Varid{b})\;\Varid{conf})\Rightarrow {}\<[E]%

@@

\>[9]{}\Varid{loop}\;(\Varid{second}\;((\Varid{mapArr}\;\Varid{rightRotate}\mathbin{>\!\!>\!\!>}\Varid{lazy})\mathbin{*\!*\!*}(\Varid{arr}\;\Varid{rightRotate}\mathbin{>\!\!>\!\!>}\Varid{lazy}))\mathbin{>\!\!>\!\!>}{}\<[E]%

@@

\>[9]{}(\Varid{arr}\;\Varid{length}\mathbin{>\!\!>\!\!>}\Varid{arr}\;\Varid{unshuffle})\mathbin{\&\!\&\!\&}(\Varid{shuffle}\mathbin{>\!\!>\!\!>}\Varid{parMap}\;\Varid{conf}\;(\Varid{ptorus}\;\Varid{f}))\mathbin{>\!\!>\!\!>}\Varid{app}\mathbin{>\!\!>\!\!>}{}\<[E]%

@@

\>[9]{}\Varid{arr}\;(\Varid{c},\Varid{a},\Varid{b})\;(\Varid{d},\Varid{a},\Varid{b})\to \Varid{arr}\;(\Varid{c},\Varid{fut}\;\Varid{a},\Varid{fut}\;\Varid{b})\;(\Varid{d},\Varid{fut}\;\Varid{a},\Varid{fut}\;\Varid{b}){}\<[E]%

@@

\>[B]{}\Varid{ptorus}\;\Varid{f}\mathrel{=}\Varid{arr}\;(\lambda \mathord{\sim}(\Varid{c},\Varid{a},\Varid{b})\to (\Varid{c},\Varid{get}\;\Varid{a},\Varid{get}\;\Varid{b}))\mathbin{>\!\!>\!\!>}\Varid{f}\mathbin{>\!\!>\!\!>}\Varid{arr}\;(\lambda \mathord{\sim}(\Varid{d},\Varid{a},\Varid{b})\to (\Varid{d},\Varid{put}\;\Varid{a},\Varid{put}\;\Varid{b})){}\<[E]%

@@

\column{30}{@{}>{\hspre}l<{\hspost}@{}}%

@@

\>[B]{}\Varid{nodefunction}\mathbin{::}\Conid{Int}\to ((\Conid{Matrix},\Conid{Matrix}),[\mskip1.5mu \Conid{Matrix}\mskip1.5mu],[\mskip1.5mu \Conid{Matrix}\mskip1.5mu])\to ([\mskip1.5mu \Conid{Matrix}\mskip1.5mu],[\mskip1.5mu \Conid{Matrix}\mskip1.5mu],[\mskip1.5mu \Conid{Matrix}\mskip1.5mu]){}\<[E]%

@@

\>[B]{}\Varid{nodefunction}\;\Varid{n}\;((\Varid{bA},\Varid{bB}),\Varid{rows},\Varid{cols})\mathrel{=}([\mskip1.5mu \Varid{bSum}\mskip1.5mu],\Varid{bA}\mathbin{:}\Varid{nextAs},\Varid{bB}\mathbin{:}\Varid{nextBs}){}\<[E]%

@@

\>[B]{}\hsindent{17}{}\<[17]%
\>[17]{}\mathbf{where}\;\Varid{bSum}\mathrel{=}\Varid{foldl'}\;\Varid{matAdd}\;(\Varid{matMult}\;\Varid{bA}\;\Varid{bB})\;(\Varid{zipWith}\;\Varid{matMult}\;\Varid{nextAs}\;\Varid{nextBs}){}\<[E]%

@@

\>[17]{}\hsindent{13}{}\<[30]%
\>[30]{}\Varid{nextAs}\mathrel{=}\Varid{take}\;(\Varid{n}\mathbin{-}\mathrm{1})\;\Varid{rows}{}\<[E]%

@@

\>[17]{}\hsindent{13}{}\<[30]%
\>[30]{}\Varid{nextBs}\mathrel{=}\Varid{take}\;(\Varid{n}\mathbin{-}\mathrm{1})\;\Varid{cols}{}\<[E]%
